Modeling the impact of IT- and AI-related emissions in a university setting

### 2025-09-11 doc/paper.tex
Change citation style for journal
Add brief discussions of Larsen and Garcia papers.

% boavizta mention immaturity of assessments and variability of estimates

\documentclass[11pt]{article}
\usepackage[margin=0.9in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{libertine}
\renewcommand{\sfdefault}{fvs} % Use as sans-serif
\renewcommand{\ttdefault}{fvm} % Use as monospace
\renewcommand{\ttfamily}{\small\fontfamily{\ttdefault}\selectfont} % Slightly smaller monospace font
\newcommand{\eg}{{\em e.g.}}
\newcommand{\coe}{CO$_2$e}
\newcommand{\gcoe}{g~\coe}
\newcommand{\tcoe}{t~\coe}
\newcommand{\kgcoe}{k\gcoe}
\newcommand{\gcoekwh}{\gcoe\ / kWh}
%\usepackage{helvet} % Helvetica font for sans-serif
%\usepackage{courier} % Courier font for monospace
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
    hidelinks
}
\usepackage[style=numeric]{biblatex}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{titling}

@@

\addbibresource{ref.bib}
\title{Modeling the impact of IT- and AI-related emissions in a university setting}
\author{Massimiliano Poletto\thanks{Independent consultant, max.poletto@gmail.com}, Nils Handler\thanks{University of Zürich, nils.handler@uzh.ch}}

@@

@@

In this study, our analysis has been bottom up, building a model of the university's IT emissions from individual estimates of devices, queries, etc. There are important tradeoffs between bottom-up precision and top-down comprehensiveness~\cite{hilty:2009:lca}. Bottom-up approaches provide higher accuracy for specific sources and transparency through clear traceability. They also make it easier to identify specific improvement areas. However, they require extensive primary data collection, require up-front decisions about what components to include or not, and may miss system-wide effects. Conversely, top-down approaches offer efficiency and a broader system perspective, but sacrifice the precision and granularity necessary for operational decision-making.

>>>>

\documentclass[11pt]{article}
\usepackage[margin=0.9in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{libertine}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
    hidelinks
}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{titling}

\usepackage[style=authoryear]{biblatex}
\addbibresource{ref.bib}
\let\cite\parencite

\renewcommand{\sfdefault}{fvs} % Use as sans-serif
\renewcommand{\ttdefault}{fvm} % Use as monospace
\renewcommand{\ttfamily}{\small\fontfamily{\ttdefault}\selectfont} % Slightly smaller monospace font
\newcommand{\eg}{{\em e.g.}}
\newcommand{\coe}{CO$_2$e}
\newcommand{\gcoe}{g~\coe}
\newcommand{\tcoe}{t~\coe}
\newcommand{\kgcoe}{k\gcoe}
\newcommand{\gcoekwh}{\gcoe\ / kWh}

@@

\title{Modeling the impact of IT- and AI-related emissions in a university setting}
\author{Massimiliano Poletto\thanks{Independent researcher, max.poletto@gmail.com}, Nils Handler\thanks{University of Zürich, nils.handler@uzh.ch}}

@@

Studies such as \textcite{larsen:2013} assess the carbon footprint of a university but do not consider IT-related emissions specifically. For example, Larsen et al. lump IT with ``equipment'', including machinery and science and technology equipment.

@@

In this study, our analysis has been bottom up, building a model of the university's IT emissions from individual estimates of devices, queries, etc. Some other studies, including \textcite{larsen:2013} and \textcite{garciaalaminos:2022}, have used top-down approaches based on environmental input-output (EIO) models. There are important tradeoffs between bottom-up precision and top-down comprehensiveness~\cite{hilty:2009:lca}. Bottom-up approaches provide higher accuracy for specific sources and transparency through clear traceability. They also make it easier to identify specific improvement areas. However, they require extensive primary data collection, force up-front decisions about what components to include or not, and may miss system-wide effects. Conversely, top-down approaches offer efficiency and a broader system perspective, but sacrifice the precision and granularity necessary for operational decision-making.
